### 2023-03-08 paper/paper-CableRoute.tex

% !TeX spellcheck = en_US

\documentclass[acmtog]{acmart}
\usepackage{amsmath}
\usepackage{multirow}
\usepackage{subfig}
\usepackage[acronym]{glossaries}
\usepackage{todonotes}
\usepackage{tikz}
\usepackage{graphicx}
\usepackage{makecell}

\AtBeginDocument{%
	\providecommand\BibTeX{{%
			Bib\TeX}}}

\setcopyright{acmcopyright}
\copyrightyear{2022}
\acmYear{2022}
\acmDOI{XXXXXXX.XXXXXXX}

%%
%% These commands are for a JOURNAL article.
\acmJournal{JACM}
\acmVolume{37}
\acmNumber{4}
\acmArticle{28}
\acmMonth{2}

\newacronym{wfs}{wfs}{Web Feature Service}
\newacronym{wps}{wps}{Web Processing Service}
\newacronym{res}{res}{resolution}
\newacronym{r}{r}{ratio}
\newacronym{al}{al}{all touched}
\newacronym{t}{t}{True}
\newacronym{f}{f}{False}
\newacronym{agg.}{agg.}{aggregated}
\newacronym{rel}{rel}{relative}
\newacronym{corr}{corr}{corrected}
\newacronym{MCDA}{MCDA}{multi-criteria decision analysis}

\begin{document}
	\title{Web Processing - Standardised GIS Analyses for Cable Route Planning}

	\author{Sebastian Heiden}
	\email{u38439@hs-harz.de}
	\affiliation{%
		\institution{Harz University of Applied Sciences}
		\streetaddress{Friedrichstrasse 57-59}
		\city{Wernigerode}
		\state{Saxony-Anhalt}
		\country{Germany}
		\postcode{38855}
	}

	\renewcommand{\shortauthors}{Heiden}

	\begin{abstract}
		The least cost paths for a power line route from a power station is computed to a power station further inland.
		Down sampling and clipping as studied as techniques to speed-up the computation.
		The path from clipped cost raster often matches the original path.
		The speed-up of the clipping can not be known before hand.
	\end{abstract}

	%%
	%% The code below is generated by the tool at https://dl.acm.org/ccs.cfm.
	%% Please copy and paste the code instead of the example below.
	%%
	\begin{CCSXML}
		<ccs2012>
		<concept>
		<concept_id>10010147.10010341</concept_id>
		<concept_desc>Computing methodologies~Modeling and simulation</concept_desc>
		<concept_significance>500</concept_significance>
		</concept>
		<concept>
		<concept_id>10002944.10011123.10011133</concept_id>
		<concept_desc>General and reference~Estimation</concept_desc>
		<concept_significance>100</concept_significance>
		</concept>
		</ccs2012>
	\end{CCSXML}

	\ccsdesc[500]{Computing methodologies~Modeling and simulation}
	\ccsdesc[100]{General and reference~Estimation}

	%%
	%% Keywords. The author(s) should pick words that accurately describe
	%% the work being presented. Separate the keywords with commas.
	\keywords{GIS, least cost path}

	\received{20 February 2007}
	\received[revised]{12 March 2009}
	\received[accepted]{5 June 2009}

	%%
	%% This command processes the author and affiliation and title
	%% information and builds the first part of the formatted document.
	\maketitle

	\section{Introduction}\label{sec:introduction}

	The German power system adds a lot of renewable energy sources in the north, because wind energy reaches highest efficiency with high wind speeds.
	At the same time old power plants e.g.\ nuclear, hard coal and lignite are going out off service~\cite{eser-frey_analyzing_2012}.
	These older power plants were mainly situated in southern and central Germany. 
	The power sink, which is industrial and private demand does not shift northwards.
	Therefore, the renewable energies have to be transported from north to south which increases the congestion in the power grid. 
	The amount of offshore wind power, that the German energy system can use, can be greatly increased by adding new power lines~\cite{leuthold_nodal_2005}.

	But for building power lines, finding the shortest path is not sufficient.
	Additional parameters also have to be taken into consideration.
	As the steepness of a road or the soil, play an important role for the building cost of a road or pipeline~\cite{suleiman_optimal_2015}.
	For planing the additional routes for a power grid, further aspects as legal regulations and acceptance
	by the local population have to be taken in consideration.
	Also technical aspects, as the effects on the grid stability are further points take into consideration~\cite{schafer_understanding_2022}.

	Due to the increasing need for renewable energy wind offshore wind turbines supply 5.5~\% annual percentage in 2020 of the German energy mix~\cite{noauthor_nettostromerzeugung_2021}.

	Other methods as the \acrfull{MCDA}  consider further aspects as stakeholder, experts, public and decision maker~\cite{bertsch_participatory_2016} (see section~\ref{sec:related-works}).

	For planning a possible route for a power line the least cost path algorithm is applied.
	The least cost path algorithm is a Dijkstra algorithm~\cite{dijkstra_note_1959} applied on a raster map.
	The vertices of the graph are the pixel centers, that are connected to the eight neighboring pixel centers via the edges.
	This enables to find routes on graphs that are not predefined like road notworks.
	The weights of the edges are the local costs to travel from one pixel center to the neighboring pixel center.
	The costs might be physical costs, as the local slope, but also can be composed on other factors as the acceptance rate for the land usage of a given class.
	The least cost path algorithm consists of at least two sequential steps.
	1) The first step is to aggregate the costs of traveling from the starting point to a given set of end points.
	This step generates the aggregated cost raster of traveling from the starting point to any point of the cost raster.
	2) In the next step the back tracking, the route of the actual least cost path is calculated.
	For each ending point the path via the lowest cost neighbor is taken until arriving at the starting point.

	Some implementations switch the role of start and ending points, so that either many starting points and one end point, or one end point and many endpoints can be used.
	In some implementation there is a extra step between 1) and 2) that calculates cost-weight direction raster, that encodes the direction of the shortest
	path to the starting point as integer values.

	\section{Methods}\label{sec:methods}
	\input{methods}

	\section{Results}\label{sec:results}
	\input{results}

	\section{Discussion}\label{sec:discussion}
	\input{discussion}

	\section{Related Works}\label{sec:related-works}

	Older models are e.g.\ are optimizing the well fare, economic modeling is not sufficient for planning modern routes, as the aspect of environmental sustainability, security of supply and the public acceptance more and more play an important role~\cite{hauff_gesellschaftliche_2011} in modern planning.

	In contrast to a GIS analysis, these further factors can be studied done with a \acrfull{MCDA}~\cite{bertsch_participatory_2016}.

	Stakeholder as decision maker can be included and combined with an expert system~\cite{bertsch_participatory_2016}.

	The uncertainties connected with \acrshort{MCDA} data uncertainties, preferential uncertainties and model uncertainties are examined by a with a sensitivity analysis~\cite{bertsch_participatory_2016} or simulation~\cite{butler_simulation_1997}.
	Both inter and intra criteria preferential uncertainties~\cite{bertsch_sensitivity_2007} can be considered.

	The different models of the power system, that are used in the expert system span from not modeling the grid at all to models that includes Kirchhoffs laws~\cite{bertsch_participatory_2016}.
	The calculation of a power system is complex, as changing one edge will change the flow in all parts of grid and the actual grid data are confidential~\cite{bertsch_participatory_2016}.
	Some models include not only the grid of a nation alone, but consider neighboring states or the overall European power system~\cite{dietrich_will_2010}.

	\section{Conclusion}\label{sec:conclusion}
	The cost of computation of the least cost path scales with the square of the resolution.
	The difference between the aggregated costs of the paths per resolution computed from the all touched false raster and the all touched true raster only shows a linear decrease.
	Hence, the gain in precision per compute time decreases.
	The presented methods tried to circumvent this bottleneck.
	When using the researched strategies, on medium resolution raster, the computation for high resolution results might successfully be reduced, without compromising the run time for worse distances to the least cost path from the high resolution raster.
	The paths might change very strongly even with small changes in total costs.
	Hence to develop an alternative backtracking algorithm that, that generates a set of possible paths within a corridor of costs might be a good strategy, to offer alternative routes and show variations between these routes at the same time.
	An alternative method for future methods, would be to specify a range of costs and compute the aggregation of the costs and superimpose these aggregations and compute the path by backtracking from the superimposed cost raster. 
	This in turn increases the computation time again.

	In an actual search for least cost path a survey might be used as method to estimate the costs.
	When the sample size of the survey is large enough, the
	weights might take local differences relative acceptance into consideration.

	Only changes in the algorithm have been applied in the search for a speed-up.
	There a still other methods as just-in-time compilation.

%%\begin{acks}
%%	$\ldots$
%%\end{acks}

%%
%% The next two lines define the bibliography style to be used, and
%% the bibliography file.
%%\bibliographystyle{ACM-Reference-Format}
\bibliographystyle{unsrtdin}
\bibliography{Web_Processing.bib}

%%
%% If your work has an appendix, this is the place to put it.
%%\appendix

%%\section{Research Methods}\label{sec:research-methods}

%%	\subsection{Part One}\label{subsec:part-one}

%%	\subsection{Part Two}\label{subsec:part-two}

%%	\section{Online Resources}\label{sec:online-resources}

\end{document}
\endinput
%%
%% End of file `sample-acmsmall.tex'.