### 2023-03-08 paper/methods.tex
grammar + change to en_GB

% !TeX spellcheck = en_GB

We retrieve a set of different spacial data-sets from  public sources as a basis for creating the cost raster.
The study area are the counties of Cuxhaven and Osterholz in the state of Lower Saxony, Germany.
Areas protected by different European and National conservation laws are provided by the German Environment Agency
as \acrfull{wfs}~\cite{noauthor_schutzgebiete_2015}.

The nationwide land coverage (ATKIS) with a scale of 1:250000 are provided by the Federal Agency for Cartography
and Geodesy~\cite{noauthor_digitales_2021}.
The national power grid (tags: 'power': line) has been retrieved via OpenStreetMap~\cite{boeing_osmnx_2017}.
Local data as houses at Level of Detail 1 are provided by the State Office for Geoinformation and Land Surveying of
Lower Saxony~\cite{noauthor_opengeodatani_2022}.
In addition, local planning geodata for the land use are taken
from 'Metropolplaner' (Planning data Lower Saxony \& Bremen)~\cite{noauthor_metropolplaner_2022}.

PyWPS~\cite{noauthor_welcome_2016} is used to provide the least cost path algorithm as a \acrfull{wps} in combination with flask~\cite{noauthor_flask_nodate}.
As client Birdy~\cite{noauthor_birdy_nodate} connects to the \acrshort{wps}, sends the cost raster and receives the resulting least cost path.
The initial implementation of the Least Cost Path algorithm is based on the implementation for the QGIS-Plugin
'Least Cost Path'~\cite{noauthor_leastcostpathdijkstra_algorithmpy_2022} in version 1.0.

The different layers of the different entities are optionally filtered, buffered and then rasterised.
Filtering the layers of the files for special attributes enables further differentiation.
For examples, it possible to differentiate between heath and uncultivated land in the land cover.
Adding a buffer can be used either used to convert a line objects such as power grids and streets into a polygon with the
correct physical width, or to add minimum distance from an existing of planed area to the new power grid.
Each of theses rasterised layers are given a weight, or cost which expresses the cost of using land covered by this layer.
The costs of all layers is aggregated with the maximum function.
Thus, an area covered by several layers is uniformly used with the highest cost.
Any place in the study area, that is not covered by any layer and thus does not yet have a weight, is given the default cost.

The costs have been grouped into different levels~(see table~\ref{tab:1}), starting from \textit{preferential} areas with
very low costs, via \textit{no restriction}s, which is the default, used when no other layers are covering the local area,
to \textit{restricted}, \textit{strongly restricted} and \textit{prohibited} areas with high costs.
These higher costs resemble the degree to which a place with this cost should be avoided, while routing the path.
The ratio of the higher costs to the lower costs directly translates into the additional detour in pixels
the algorithm is willing to go.
Thus, as \textit{prohibited} areas describe a legal obligation, not to use these areas or only to the utmost minimum,
the weight that resembles the costs for these types of areas, has to be especially high.\\

All these layers are provided as vectors.
The Least Cost Path algorithm uses raster data.
Rasterisation transforms a vector into a raster.
Rasterisation can be done in two different ways.
In both ways, the rasterisation can be imaged, as the old vector is superimposed on the new raster grid with the new
given resolution and the new affine transformation and the coordinate reference system of the vector.
Both rasterisation techniques differ in the selection of the pixel, that describe the original polygon.
The pixels can be selected , if either the centre of that pixel is overlapped
by the geo-object, or any part of the pixel is overlaid.
The version with any part overlapped is called all touched True.
The version where a overlapped pixel centre is require is called all touched False.
All touched False is considered the default (see figure~\ref{fig:alltouched}).
\begin{figure}[!ht]
	\centering

	\resizebox{8cm}{!}{
		\begin{tikzpicture}
			\node[draw] at (0,6.5) {\huge all touched false};
			\draw[step=1cm,gray,very thin] (-1.9,-1.9) grid (5.9,5.9);
			\draw[fill=blue, thick] (-0.2, 0.2)--(-0.6, 3.1)--(-0.2,4.1)--(1.2,3.4)--(2.8,3.6)--(3.2, 2.6)--(3.4, 2.1)--(1.8,1.3)--(1.2, 0.1)--cycle;

			\draw[ultra thick,->, red] (6.5, 2) --node[above=1mm] {\huge rasterization} (8.5, 2);

			\draw[step=1cm,gray,very thin] (9.1,-1.9) grid (16.9,5.9);
			\draw[fill=blue] (11,0)--(11,2)--(10,2)--(10,3)--(11,3)--(11,4)--(12,4)--(12,3)--(13,3)--(13,4)--(14,4)--(14,2)--(13,2)--(13,1)--(12,1)--(12,0)--cycle;
		\end{tikzpicture}%
	}%
	\vspace*{5mm}
	\resizebox{8cm}{!}{
		\begin{tikzpicture}
			\node[draw] at (0,6.5) {\huge all touched true};
			\draw[step=1cm,gray,very thin] (-1.9,-1.9) grid (5.9,5.9);
			\draw[fill=blue, thick] (-0.2, 0.2)--(-0.6, 3.1)--(-0.2,4.1)--(1.2,3.4)--(2.8,3.6)--(3.2, 2.6)--(3.4, 2.1)--(1.8,1.3)--(1.2, 0.1)--cycle;

			\draw[ultra thick,->, red] (6.5, 2) --node[above=1mm] {\huge rasterization} (8.5, 2);

			\draw[step=1cm,gray,very thin] (9.1,-1.9) grid (16.9,5.9);
			\draw[fill=blue] (10,0)--(10,5)--(11,5)--(11,4)--(15,4)--(15,1)--(13,1)--(13,0)--cycle;
		\end{tikzpicture}%
	}%
	\caption{Graphical example for rastering a vector (left blue), to a raster (right blue) with either all-touched False (above), or all touched True (below).}
	\label{fig:alltouched}

\end{figure}

\begin{table}[h!]
	\caption{Used levels of costs, the applied numerical equivalent and example layer this cost have been used for.}
	\label{tab:1}
	\centering
	\begin{tabular}{ l  r  l }
		Cost Level 			& Cost 					& Example\\
		\hline
		Prohibited 			& 500					& \makecell[lt]{Conversation areas as\\ National Parks, Buildings} \\
		strongly Restricted & 10 					& \makecell[lt]{Conversation areas as Bird Reserve} \\
		Restricted 			& 5						& \makecell[lt]{Protected Landscape Area,\\ Industrial Areas, motorway, railway} \\
		No Restriction 		& 0.5					& Default\\
		Preferential 		& 0.1					& \makecell[lt]{Power Grid,\\ Motorway and Railway Buffers}\\
	\end{tabular}
\end{table}

The complete list of layers and their processing, can be found in \textbf{Supplement S1}.

All three steps of the generation of the least cost path: generation of the cost raster, aggregation and backtracking is shown with an example for a starting cost raster of 50~m resolution and a rasterisation with all touched False (see \ref{fig:costs2path}).

The chosen implementation applies early stopping.
Therefore, the costs for points that are not needed to try to connect to the end point are not aggregated.
After finding an aggregated cost for every end point, the aggregation stops and the backtracking starts.
Because the path ends at a power transformer, which is a building type, the paths ends at in a \textit{Prohibited} area.
Therefore, areas even further away from the starting point have been explored first.

\begin{figure}
	\centering

	\subfloat[\centering Cost Raster.]{{\includegraphics[width=.30\linewidth]{./images/CostRasterExample_cut.png} }}%
	\enskip
	\subfloat[\centering Aggregated Cost Raster.]{{\includegraphics[width=.30\linewidth]{./images/AggregatedCosts_cut.png} }}%
	\enskip
	\subfloat[\centering Least Cost Path.]{{\includegraphics[width=.30\linewidth]{./images/LeastCostPathExample_cut.png} }}%

	\caption{Figures of Cost raster and the resulting aggregated Costs and the least cost path for a resolution of 50~m.}
	\label{fig:costs2path}
\end{figure}

%% Aggregation is implemented in a priority queue. 
%% Hence a point with a lower cost is evaluated earlier as a starting position to evaluate the cost of its neighbours, than a point with higher cost.