### 2023-03-11 paper/paper-CableRoute.tex

% !TeX spellcheck = en_GB

\documentclass[acmtog]{acmart}
\usepackage{amsmath}
\usepackage{multirow}
\usepackage{subfig}
\usepackage[acronym]{glossaries}
\usepackage{todonotes}
\usepackage{tikz}
\usepackage{graphicx}
\usepackage{makecell}

\AtBeginDocument{%
	\providecommand\BibTeX{{%
			Bib\TeX}}}

\setcopyright{acmcopyright}
\copyrightyear{2022}
\acmYear{2022}
\acmDOI{XXXXXXX.XXXXXXX}

%%
%% These commands are for a JOURNAL article.
\acmJournal{JACM}
\acmVolume{37}
\acmNumber{4}
\acmArticle{28}
\acmMonth{2}

\newacronym{wfs}{wfs}{Web Feature Service}
\newacronym{wps}{wps}{Web Processing Service}
\newacronym{MCDA}{MCDA}{multi-criteria decision analysis}

\begin{document}
	\title{Web Processing - Standardised GIS Analyses for Cable Route Planning}

	\author{Sebastian Heiden}
	\email{u38439@hs-harz.de}
	\affiliation{%
		\institution{Harz University of Applied Sciences}
		\streetaddress{Friedrichstrasse 57-59}
		\city{Wernigerode}
		\state{Saxony-Anhalt}
		\country{Germany}
		\postcode{38855}
	}

	\renewcommand{\shortauthors}{Heiden}

	\begin{abstract}
		The least cost paths for a power line from a power station at the cost to a power station further inland are calculated.
		Down-sampling and clipping are investigated as techniques to speed up the calculation.
		The path from the clipped cost raster is often the same as the original path.
		The speed-up of the clipping can not be known in advance.
	\end{abstract}

	%%
	%% The code below is generated by the tool at https://dl.acm.org/ccs.cfm.
	%% Please copy and paste the code instead of the example below.
	%%
	\begin{CCSXML}
		<ccs2012>
		<concept>
		<concept_id>10010147.10010341</concept_id>
		<concept_desc>Computing methodologies~Modeling and simulation</concept_desc>
		<concept_significance>500</concept_significance>
		</concept>
		<concept>
		<concept_id>10002944.10011123.10011133</concept_id>
		<concept_desc>General and reference~Estimation</concept_desc>
		<concept_significance>100</concept_significance>
		</concept>
		</ccs2012>
	\end{CCSXML}

	\ccsdesc[500]{Computing methodologies~Modeling and simulation}
	\ccsdesc[100]{General and reference~Estimation}

	%%
	%% Keywords. The author(s) should pick words that accurately describe
	%% the work being presented. Separate the keywords with commas.
	\keywords{GIS, least cost path}

	\received{14 March 2023}
	%%\received[revised]{12 March 2009}
	%%\received[accepted]{5 June 2009}

	%%
	%% This command processes the author and affiliation and title
	%% information and builds the first part of the formatted document.
	\maketitle

	\section{Introduction}\label{sec:introduction}

	The German power system is adding a lot of renewable energy sources in the north, because wind energy reaches its highest efficiency with high wind speeds.
	At the same time old power plants e.g.\ nuclear, hard coal and lignite are being phased out~\cite{eser-frey_analyzing_2012}.
	These older power plants were mainly located in southern and central Germany. 
	The energy sink,  industrial and private demand, is not shifting north.
	Therefore, the renewable energy has to be transported from north to south which increases the congestion in the power grid. 
	The amount of offshore wind power, that the German energy system can use, can be greatly increased by adding new power lines~\cite{leuthold_nodal_2005}.

	However, it is not enough to find the shortest route when building new power lines.
	Other parameters have to be taken into consideration.
	As the steepness of a road or the soil, play an important role for the building cost of a road or pipeline~\cite{suleiman_optimal_2015}.
	When planing the additional routes for a power grid, further aspects such as legal regulations and acceptance
	by the local population have to be taken in consideration.
	Also technical aspects, as the effects on the grid stability are further points take into consideration~\cite{schafer_understanding_2022}.

	Due to the increasing demand for renewable energy wind offshore wind turbines supply 5.5~\% annual percentage in 2020 of the German energy mix~\cite{noauthor_nettostromerzeugung_2021}.

	Other methods such as \acrfull{MCDA} consider further aspects as stakeholder, experts, public and decision maker~\cite{bertsch_participatory_2016} (see section~\ref{sec:related-works}).

	The least cost path algorithm is used to plan a possible route for a power line.
	The least cost path algorithm is a Dijkstra algorithm~\cite{dijkstra_note_1959} applied on a raster map.
	The vertices of the graph are the pixel centres, that are connected to the eight neighbouring pixel centres via the edges.
	This makes it possible to find routes on graphs that are not predefined, such as road networks.
	The weights of the edges are the local costs of travel from one pixel center to the neighbouring pixel center.
	The costs can be physical costs, such as the local slope, but also can be composed of other factors as the acceptance rate for the land usage of a given class.
	The Least Cost Path algorithm consists of at least two successive steps.
	1) The first step is to aggregate the costs of travelling from the starting point to a given set of end points.
	This step generates the aggregated cost raster of travelling from the starting point to any point of the cost raster.
	2) In the next step the backtracking, the route of the actual Least Cost Path is calculated.
	For each end point the path via the lowest cost neighbour is taken until the starting point is reached.

	Some implementations switch the role of start and ending points, so that either many start points and one single end point, or one single end point and many end points can be used.
	In some implementation there is an extra step between 1) and 2) that calculates cost-weight direction raster, that encodes the direction of the shortest
	path to the starting point as integer values.

	\section{Methods}\label{sec:methods}
	\input{methods}

	\section{Results}\label{sec:results}
	\input{results}

	\section{Discussion}\label{sec:discussion}
	\input{discussion}

	\section{Related Works}\label{sec:related-works}

	The different models of the power system, that are used in the expert system range from not modelling the grid at all to models that include Kirchhoffs laws~\cite{bertsch_participatory_2016}.
	The calculation of a power system is complex, because changing one edge changes the flow in all parts of grid and the actual grid data are confidential~\cite{bertsch_participatory_2016}.
	Some models consider not only the grid of one nation, but consider neighbouring states or the whole European power system~\cite{dietrich_will_2010}.	
	Older national power grid models are e.g.\ are optimising the well fare, but economic modelling is not sufficient for planning modern routes, as the aspect of environmental sustainability, security of supply and the public acceptance play an increasingly important role~\cite{hauff_gesellschaftliche_2011} in modern planning.

	In contrast to a GIS analysis, these other factors can be studied done with a \acrfull{MCDA}~\cite{bertsch_participatory_2016}.
	Stakeholders as decision makers can be included and combined with an expert system~\cite{bertsch_participatory_2016}.

	The uncertainties associated with \acrshort{MCDA} data uncertainties, preferential uncertainties and model uncertainties are investigated by a with a sensitivity analysis~\cite{bertsch_participatory_2016} or simulation~\cite{butler_simulation_1997}.
	Both inter- and intra-criteria preferential uncertainties~\cite{bertsch_sensitivity_2007} can be considered.

	\section{Conclusion}\label{sec:conclusion}
	The cost of computing of the Least Cost Path scales with the square of the resolution.
	The difference between the aggregated costs of the paths per resolution computed from the all touched false raster and the all touched true raster only shows a linear decrease.
	Therefore, the gain in accuracy per compute time decreases.
	The presented methods have attempted to circumvent this bottleneck.
	When these strategies are applied to a medium resolution raster, the compute time for high resolution results can successfully be reduced, without compromising the run time for worse distances to the least cost path from the high resolution raster.
	The paths can change significantly even for a small change in total cost.
	Therefore, to develop an alternative backtracking algorithm, that generates a corridor of costs might be a good strategy, to offer alternative paths and at the same time show variations between these paths.
	An alternative for future methods, would be to specify a range of costs and compute the aggregation of the costs and superimpose these aggregations and compute the path by backtracking from the superimposed cost raster. 
	This again increases the computation time.

	In an actual search for Least Cost Path search, a survey could be used as a method to estimate the costs.
	If the sample size of the survey is large enough, the
	weights could take  into account local differences in relative acceptance.

	Only changes in the algorithm have been applied in the search for acceleration.
	Therefore, other methods such as just-in-time compilation have not been tested.

%%\begin{acks}
%%	$\ldots$
%%\end{acks}

%%
%% The next two lines define the bibliography style to be used, and
%% the bibliography file.
%%\bibliographystyle{ACM-Reference-Format}
\bibliographystyle{unsrtdin}
\bibliography{Web_Processing.bib}

%%
%% If your work has an appendix, this is the place to put it.
%%\appendix

%%\section{Research Methods}\label{sec:research-methods}

%%	\subsection{Part One}\label{subsec:part-one}

%%	\subsection{Part Two}\label{subsec:part-two}

%%	\section{Online Resources}\label{sec:online-resources}

\end{document}
\endinput
%%
%% End of file `sample-acmsmall.tex'.