### 2023-03-12 paper/paper-CableRoute.tex

% !TeX spellcheck = en_GB

\documentclass[acmtog]{acmart}
\usepackage{amsmath}
\usepackage{multirow}
\usepackage{subfig}
\usepackage[acronym]{glossaries}
\usepackage{todonotes}
\usepackage{tikz}
\usepackage{graphicx}
\usepackage{makecell}

\AtBeginDocument{%
	\providecommand\BibTeX{{%
			Bib\TeX}}}

\setcopyright{acmcopyright}
\copyrightyear{2022}
\acmYear{2022}
\acmDOI{XXXXXXX.XXXXXXX}

%%
%% These commands are for a JOURNAL article.
\acmJournal{JACM}
\acmVolume{37}
\acmNumber{4}
\acmArticle{28}
\acmMonth{2}

\newacronym{wfs}{wfs}{Web Feature Service}
\newacronym{wps}{wps}{Web Processing Service}
\newacronym{MCDA}{MCDA}{multi-criteria decision analysis}

\begin{document}
	\title{Web Processing - Standardised GIS Analyses for Cable Route Planning}

	\author{Sebastian Heiden}
	\email{u38439@hs-harz.de}
	\affiliation{%
		\institution{Harz University of Applied Sciences}
		\streetaddress{Friedrichstrasse 57-59}
		\city{Wernigerode}
		\state{Saxony-Anhalt}
		\country{Germany}
		\postcode{38855}
	}

	\renewcommand{\shortauthors}{Heiden}

	\begin{abstract}
		Least Cost Paths for a power line from a power station at the shore line to a power station further inland are calculated.
		Downsampling and clipping are investigated as techniques to speed up the calculation.
		The paths from the clipped cost raster often are the same as the original paths.
		The speed-up of the clipping can not be known in advance.
	\end{abstract}

	%%
	%% The code below is generated by the tool at https://dl.acm.org/ccs.cfm.
	%% Please copy and paste the code instead of the example below.
	%%
	\begin{CCSXML}
		<ccs2012>
		<concept>
		<concept_id>10010147.10010341</concept_id>
		<concept_desc>Computing methodologies~Modeling and simulation</concept_desc>
		<concept_significance>500</concept_significance>
		</concept>
		<concept>
		<concept_id>10002944.10011123.10011133</concept_id>
		<concept_desc>General and reference~Estimation</concept_desc>
		<concept_significance>100</concept_significance>
		</concept>
		</ccs2012>
	\end{CCSXML}

	\ccsdesc[500]{Computing methodologies~Modeling and simulation}
	\ccsdesc[100]{General and reference~Estimation}

	%%
	%% Keywords. The author(s) should pick words that accurately describe
	%% the work being presented. Separate the keywords with commas.
	\keywords{GIS, Least Cost Path}

	\received{14 March 2023}
	%%\received[revised]{12 March 2009}
	%%\received[accepted]{5 June 2009}

	%%
	%% This command processes the author and affiliation and title
	%% information and builds the first part of the formatted document.
	\maketitle

	\section{Introduction}\label{sec:introduction}

	The German power system is adding renewable energy sources in the north, where wind energy plants reach their highest efficiency, due to higher wind speeds.
	At the same time old power plants e.g.\ nuclear, hard coal and lignite are being phased out~\cite{eser-frey_analyzing_2012}.
	These older power plants were mainly located in southern and central Germany. 
	The energy sink,  industrial and private demand, is not shifting north.
	Therefore, the renewable energy has to be transported from north to south which increases the congestion in the power grid. 
	The amount of offshore wind power, that the German energy system can use, can be greatly increased by adding new power lines~\cite{leuthold_nodal_2005}.

	However, it is not enough to find the shortest route when building new power lines.
	Other parameters have to be taken into consideration, as the steepness of a road or the soil for example, play an important role for the building cost of a road or pipeline~\cite{suleiman_optimal_2015}.
	When planing the additional routes for a power grid, further aspects such as legal regulations and acceptance
	by the local population have to kept in mind.
	Also technical aspects, as the effects on the grid stability are further points to take into consideration~\cite{schafer_understanding_2022}.

	Due to the increasing demand for renewable energy wind offshore wind turbines supplied 5.5~\% annual percentage in 2020 of the German energy mix~\cite{noauthor_nettostromerzeugung_2021}.

	 The physical modulation in the power systems models, range from not modelling the grid at all to models  The use of Kirchhoffs laws~\cite{bertsch_participatory_2016}.
	The calculation of a power system is complex, because changing one edge changes the flow in all parts of grid.
	Besides, the actual grid data are confidential~\cite{bertsch_participatory_2016}.
	Some models not only consider the grid of one nation, but consider neighbouring states or the whole European power system~\cite{dietrich_will_2010}.	
	Older power grid models are e.g.\ optimising the well fare, but economic modelling is not sufficient for planning modern routes, as the aspect of environmental sustainability, security of supply and the public acceptance play an increasingly important role~\cite{hauff_gesellschaftliche_2011} in modern planning.

	In contrast to a GIS analysis, these other factors can be studied done with a \acrfull{MCDA}~\cite{bertsch_participatory_2016}.
	Stakeholders as decision makers can be included and combined with an expert system~\cite{bertsch_participatory_2016}.

	The uncertainties associated with \acrshort{MCDA} are data uncertainties, preferential uncertainties and model uncertainties are investigated with a sensitivity analysis~\cite{bertsch_participatory_2016} or simulation~\cite{butler_simulation_1997}.
	Both inter- and intra-criteria preferential uncertainties~\cite{bertsch_sensitivity_2007} can be considered.

	\section{Methods}\label{sec:methods}
	\input{methods}

	\section{Results}\label{sec:results}
	\input{results}

	\section{Discussion}\label{sec:discussion}
	\input{discussion}

	\section{Conclusion}\label{sec:conclusion}
	The cost of computing of the Least Cost Path scales with the square of the resolution.
	The difference between the aggregated costs of the paths per resolution computed from the all~touched set to False raster True raster only shows a linear decrease.
	Therefore, the gain in accuracy per compute time decreases.
	The presented methods attempt to circumvent this bottleneck.
	When these strategies are applied to a medium resolution raster, the compute time for high resolution results can successfully be reduced, without compromising the run time for worse distances to the Least Cost Path from the high resolution raster.
	The paths can change significantly even for a small change in total cost.
	Therefore, to develop an alternative backtracking algorithm, that generates a corridor of costs might be a good strategy, to offer alternative paths and at the same time show variations between these paths.
	An alternative would be to specify a range of costs and compute the aggregation of the costs and superimpose these aggregations and compute the path by backtracking from the superimposed cost raster. 
	This again increases the computation time.

	In an actual search for Least Cost Path search, a survey could be used as a method to estimate the costs.
	If the sample size of the survey is large enough, the
	weights could take  into account local differences in relative acceptance.

	Only changes in the algorithm have been applied in the search for acceleration.
	Therefore, other methods such as just-in-time compilation have not been tested.

%%\begin{acks}
%%	$\ldots$
%%\end{acks}

%%
%% The next two lines define the bibliography style to be used, and
%% the bibliography file.
%%\bibliographystyle{ACM-Reference-Format}
\bibliographystyle{unsrtdin}
\bibliography{Web_Processing.bib}

%%
%% If your work has an appendix, this is the place to put it.
%%\appendix

%%\section{Research Methods}\label{sec:research-methods}

%%	\subsection{Part One}\label{subsec:part-one}

%%	\subsection{Part Two}\label{subsec:part-two}

%%	\section{Online Resources}\label{sec:online-resources}

\end{document}
\endinput
%%
%% End of file `sample-acmsmall.tex'.