### 2023-03-21 presentations/20230323_seh_final.tex
added new path for images for the presentations.
changed image paths for the 20121014_seh_kickoff.tex and 20221202_seh_midterm.tex

added new presentation 20230323_seh_final.tex with first slides.

% !TeX spellcheck = en_GB
\documentclass[usenames,dvipsnames,aspectratio=169]{beamer}
\usetheme{Szeged}
\usecolortheme{beaver}
\usepackage{tikz}
\usepackage{subfig}
\setbeamertemplate{navigation symbols}{}
\usepackage{xcolor}

\title{Web Processing - Standardized GIS Analyses for Cable Route Planning}
\author{Sebastian Heiden}
\institute{Harz University of Applied Sciences}
\date{December 02, 2022}
\begin{document}

	\begin{frame}[plain]
		\includegraphics[scale=0.21]{images/3-HSH-Logo-RGB-en.png}
		\maketitle
	\end{frame}

	\section{Review \& Topics}
	\begin{frame}{Review \& Topics}

		\begin{minipage}[t]{0.48\textwidth}
			Review
			\begin{itemize}
				\item Rasterisation
				\item Cost Raster
				\item Cost Paths
				\item PyWPS (Start)
			\end{itemize}
		\end{minipage}
		\hfill	
		\begin{minipage}[t]{0.48\textwidth}
			Topics
			\begin{itemize}
				\item Web Processing Service (Concluded)
				\item From Cost Raster to Least Cost Path
				\item Comparing Paths
				\item Speed-Up
				\begin{itemize}
					\item Clipping
					\item Down Sampling
					\item Super Position Medium Resolution
					\item Validation
				\end{itemize}

			\end{itemize}
		\end{minipage}

	\end{frame}

	\section{Web Processing Service (Concluded)}
	\begin{frame}{Web Processing Service (Concluded)}
		\begin{itemize}
		\item goals: 
		\begin{itemize}
			\item use wps as a simple, standardized way
			\item optimize search algorithm
		\end{itemize}
		\item current: 
		\begin{itemize}
			\item testing PyWPS\footnote{https://pywps.readthedocs.io/en/latest/index.html}
			\item cost path (open Dijkstra implementation - QGIS-plugin)\footnote{https://github.com/Gooong/LeastCostPath}
		\end{itemize}
	\end{itemize}
	\end{frame}

	\section{From Cost Raster to Least Cost Path}
	\begin{frame}{From Cost Raster to Least Cost Path}
		\begin{figure}
			\centering
			\subfloat[\centering Cost Raster.]{{\includegraphics[width=.30\linewidth]{images/maps/CostRasterExample_cut.png} }}%
			\enskip
			\subfloat[\centering Aggregated Cost Raster.]{{\includegraphics[width=.30\linewidth]{images/maps/AggregatedCosts_cut.png} \label{fig:aggregation}}}%
			\enskip
			\subfloat[\centering Least Cost Path.]{{\includegraphics[width=.30\linewidth]{images/maps/LeastCostPathExample_cut.png} }}%

			\caption{Figures of the cost raster and the resulting aggregated costs and the Least Cost Path for a resolution of 50~m, all~touched set to False.}
			\label{fig:costs2path}
		\end{figure}
	\end{frame}

	\section{Comparing Paths}
	\begin{frame}{Comparing: Map}
		\begin{figure}
			\centering

			\subfloat[\centering all touched False.]{{\includegraphics[width=.45\linewidth]{images/maps/LeastCostPaths_al_F_v2_small.png} }}%
			\qquad
			\subfloat[\centering all touched True.]{{\includegraphics[width=.45\linewidth]{images/maps/LeastCostPaths_al_T_v2_small.png} }}%

			\caption{Figures of the Least Cost Paths, depending on the parameter all~touched. All~touched False: dashed lines, True: continuous lines. Higher resolutions: green, lower resolutions: red.}
			\label{fig:paths_alltouched}
		\end{figure}
	\end{frame}

	\begin{frame}{Comparing: Costs}
		\begin{table}[t]
			\caption{Least Cost Paths: length~(l) for the different resolutions~(res), including the mean minimum distance~($d_{mean}$), agg. costs and the agg. costs per resolution.} 
			\label{tab:2}
			\centering
			\begin{tabular}{ r  r  r  r  r  r   }
				res /m & $l_{al=f} /m$ & $l_{al=t} /m$ & $d_{mean}$ /m  & agg. $costs_{al=f} \times m$ & agg. $costs_{al=t} \times m$ \\
				\hline
				5 	& 76136.3	& 78002.0 &  126.0  & 93329.6 &  97584.8 \\
				10 	& 75430.1 	& 77936.6 &  277.9  & 89312.5 &  97311.8 \\
				25 	& 75422.9 	& 78422.9 &  313.8  & 83871.7 &  96816.4 \\
				50 	& 76135.0	& 70620.0 & 1140.0  & 70451.2 & 115003.7 \\
				100 & 76283.8	& 74120.7 & 1946.4  & 64051.6 & 167226.8 \\
			\end{tabular}
		\end{table}
	\end{frame}

\section{Speed-Up}
	\begin{frame}{Speed-Up: Clipping}
		\begin{figure}
			\centering

			\subfloat[\centering 50~m Resolution.]{{\includegraphics[width=.45\linewidth]{images/maps/LeastCostPath_50m_small.png} }}%
			\qquad
			\subfloat[\centering 5~m, Clipped.]{{\includegraphics[width=.45\linewidth]{images/maps/LeastCostPath_5m_false_clipped_small.png} }}%

			\caption{Figures of the Least Cost Paths, on the cost raster with 50~m resolution and the computed paths, used to clip the high resolution raster.}
			\label{fig:clipping}
		\end{figure}
	\end{frame}

	\begin{frame}{Speed-Up: Down Sampling}
	\end{frame}

	\begin{frame}{Speed-Up:Super Position Medium Resolution}
	\end{frame}

	\begin{frame}{Speed-Up:Validation}
	\end{frame}
\end{document}